Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-04-22 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
small changes

As we have seen in Sec.~\ref{sec:review_theories}, QNMs of slowly-rotating
BHs in modified gravity theories can belong to two families depending on how
they behave under a parity transformation: axial and polar. Which one do
we use to match with Eqs.~\eqref{eq:qnm_nongr_iso}?
%
To answer this question one has to work with a chosen theory and perform a
translation between the metric perturbations $h_{\mu\nu}$ in the
Regge-Wheeler-Zerilli gauge~\cite{Regge:1957td,Zerilli:1970se} and connect it with the transverse-traceless gauge
used to described GWs (see e.g.~Ref.~\cite{Maggiore:2018sht}, Chapter~12),
%
In GR, both axial and polar QNMs are isospectral and hence which QNM we use to
model the ringdown makes no difference.
%
In beyond-GR theories, isospectrality is in general broken (see
in Ref.~\cite{Hui:2021cpm} for a counterexample).
%
Thus, how axial and polar gravitational QNMs appear in the GW signal has to be
answered on a theory-by-theory basis.
%
This is outside the scope of this paper and here we take the more pragmatic
approach of simply choosing the \emph{least damped gravitational mode between
the two parities.}
%
The justification is that this is the mode (if excited during the merger) which is
the most likely to appear in the signal.
We performed the mapping between theory-specific QNM calculations and the ParSpec framework
under the hypothesis above, for the theories listed in Sec.~\ref{sec:review_theories}.
%
We summarize our results in Table~\ref{tab:ref_theories_qnms} and
leave the details of our calculations to Appendix~\ref{app:map_details}.

@@

Let us discuss some avenues for future work.
%
First, we could implement a high-spin version of the GR fitting coefficients to the ParSpec formulas.
%
This has already been done in Ref.~\cite{Carullo:2021dui} extending the validity of the ParSpec formulas up to
spins of $\chi_{\rm f} \approx 0.99$.
%
For the events analyzed here, the original fit by Ref.~\cite{Maselli:2019mjd}
was sufficient, but it might not be the case with upcoming GW observation campaigns.
%
Second, it would be important to incorporate additional effects, such as spin-precession and eccentricity
to \pSEOB~(see e.g., Refs~\cite{Ossokine:2020kjp,Ramos-Buades:2021adz}).

>>>>

As we have seen in Sec.~\ref{sec:review_theories}, QNMs of slowly-rotating
BHs in modified gravity theories can belong to two families depending on how
they behave under a parity transformation: axial and polar. Which one do
we use to match with Eqs.~\eqref{eq:qnm_nongr_iso}?
%
To answer this question one has to work with a chosen theory and perform a
translation between the metric perturbations $h_{\mu\nu}$ in the
Regge-Wheeler-Zerilli gauge~\cite{Regge:1957td,Zerilli:1970se} and connect it with the transverse-traceless gauge
used to described GWs (see, for instance,~Ref.~\cite{Maggiore:2018sht}, Chapter~12),
%
In GR, both axial and polar QNMs are isospectral and hence which QNM we use to
model the ringdown makes no difference.
%
In beyond-GR theories, isospectrality is in general broken (see
in Ref.~\cite{Hui:2021cpm} for a counterexample).
%
Thus, how axial and polar gravitational QNMs appear in the GW signal has to be
answered on a theory-by-theory basis.
%
This is outside the scope of this paper and here we take the more pragmatic
approach of simply choosing the \emph{least damped gravitational mode between
the two parities.}
%
The justification is that this is the mode (if excited during the merger) which is
the most likely to appear in the signal.
We performed the mapping between theory-specific QNM calculations and the ParSpec framework
under the hypothesis above, for the theories listed in Sec.~\ref{sec:review_theories}.
%
We summarize our results in Table~\ref{tab:ref_theories_qnms} and
leave the details of our calculations to Appendix~\ref{app:map_details}.

@@

Let us discuss some avenues for future work.
%
First, we could implement a high-spin version of the GR fitting coefficients to the ParSpec formulas.
%
This has already been done in Ref.~\cite{Carullo:2021dui} extending the validity of the ParSpec formulas up to
spins of $\chi_{\rm f} \approx 0.99$.
%
For the events analyzed here, the original fit by Ref.~\cite{Maselli:2019mjd}
was sufficient, but it might not be the case with upcoming GW observation campaigns.
%
Second, it would be important to incorporate additional effects, such as spin-precession and eccentricity
to \pSEOB~(see, for instance, Refs~\cite{Ossokine:2020kjp,Ramos-Buades:2021adz}).
